Automated Topic Naming

### 2012-03-15 ese/ese.tex
updated todos for 'minor changes'
minor fixes

% migod
In general, the fruits of mining software artifacts are often project
specific and hard to generalize.  
However, in our previous work we
investigated \emph{topic trends} --- that is, topics that recur over
time --- we observed that topic trends often corresponded to
\emph{non-functional requirements} (NFRs)~\cite{Hindle09ICSM}, which is
further emphasized in this paper due to the large numbers of NFR
labelled topics.  
This is encouraging, as NFRs have the property of being cross-domain
and widely applicable. 
In this sense, they are useful abstractions for developer
conversations about different software projects.  
Furthermore, there is a series of standards on NFRs, such as ISO9126 \cite{iso9126}, that are specifically intended to apply to projects of varying
types; this suggests that our goal of trying to extract NFR-related development topics, such as those related to software quality models, holds promise.

@@

%We first introduces labelled topic extraction, and then uses that technique to analyze development activity in two large-scale examples.
We begin by discussing related work in Section \ref{sec:related}.
Next, we describe how we generated our data (Section \ref{sec:wordlist}). For semi-unsupervised classification (Section \ref{sec:unsuplabelling}), we
begin by creating word-lists to signify when a topic matches an NFR label. We then apply our classifier and analyze the results. 
%and how we created the word-lists used in unsupervised classification. 
%We show that labels with their topics can be learned and used to classify other data-sets, either without training 
%Our goal is to label a given topic with either an NFR from our taxonomy,
 %Section \ref{sec:unsuplabelling} presents our technique for using word-lists to do unsupervised classification, and describes the results.
In Section \ref{sec:suplabelling}, we manually annotate the topics, and use those annotations as training data for supervised classification.  
%We then present visualizations of named topics and their trends over time to aid communication and analysis. 
To demonstrate an application of labelled topic extraction, we use an exploratory case study of three open source database systems to show how named
topics can be compared between projects  (Section \ref{sec:analysis}). 
The paper concludes with a discussion of limitations (Section \ref{sec:limit}), and future work.

@@

With ConcernLines, Treude et al.~\cite{treude09cl} showed tag occurrence
using colour and intensity, and our plots that rely on color and intensity
have a similar look and feel to Treude et al.'s plots.
They mined
developer created 
change request tags from IBM Jazz
repositories and used these 
to analyze the evolution of a single product.
Change requests in Jazz allow for users to annotate each change
request and check-in with tags such as ``ui'', ``usability'' or
``milestone 3''. 
% make evolutionary analyses of a single product. 
The presence of a well-maintained set of tags is obviously essential to the success of this technique.

@@

This paper and our MSR 2011 paper~\cite{msr2011} are based off of the
work of Ernst and Mylopoulos~\cite{ernst10refsq} and the work of Hindle et
al.~\cite{Hindle09ICSM}.
In Ernst and Mylopoulos~\cite{ernst10refsq}, we describe an earlier
project that identifies changes in quality requirements
in GNOME software projects (GNOME is a Linux desktop environment). Unlike this paper, this earlier approach was more exploratory and had
less validation. In particular, it  uses different word-lists, solely uses text-matching,
and does not leverage machine learning strategies.  
Our approach in this paper, and our
MSR 2011 paper~\cite{msr2011}, extends Ernst and Mylopoulos~\cite{ernst10refsq} by using 
word-lists to label topics, which are completely absent in the earlier work.
 Hindle et
al.~\cite{Hindle09ICSM}
 propose a windowed method of topic
analysis that we extend with labelled
topics, NFRs and new visualizations. This windowed method was
to bucket documents  by time windows (such as months), that could overlap if
necessary, and then use LDA to extract documents from these
bucket. These topics were then related to topics in other adjacent
buckets, and joined if deemed sufficiently similar. This paper, and our
MSR'2011 paper~\cite{msr2011}, extend the method of Hindle et
al.~\cite{Hindle09ICSM} by labelling topics
and providing an alternative visualization, the topic time-line
described in Section \ref{sec:analysis}.

@@

For each project, we studied source control commit comments, the messages
that programmers write when they commit revisions to a source control
repository. 
Most commits we observed had comments: 90\% in MySQL 3.23,
98.5\% in PostgreSQL and 99.99\% in MaxDB 7.500.
Commit comments are often studied by researchers, as they are the most readily accessible source of project interactions, and developers are often
required to create them by the repository mechanism (e.g., CVS).  Additionally, relying only on commit comments makes our approach more generalizable,
as we do not assume the presence of other artifact corpora.
%This is the most readily accessible source of project interactions for outside researchers, since developers often or always (e.g., Git) write such
%message. Our approach is more generalizable if it does not assume the presence of other corpora. 
%bug trackers and email-list archives were not available for both projects. 
An example of a typical commit message, from MySQL, is: \textit{``history annotate diffs bug fixed (if mysql\-\_real\-\_connect() failed there were
two pointers to malloc'ed strings, with memory corruption on free(), of course)''}. 
We extracted these messages and indexed them by creation time. 
Each word in the message was
stripped of punctuation and converted to lowercase.
We summarized each message as a word distribution minus stop-words
such as ``\emph{the}'' and ``\emph{at}''. We did not stem or apply any
other transforms to the messages.
Our stop words are derived from the Natural Language Toolkit (NLTK) English stop-word
list~\footnote{NLTK: \url{http://www.nltk.org/}}.

@@

%For MySQL (Figure \ref{fig:mysql-timeline}), we similarly validated our NFR patterns with external mailing list data. 
The development period of MySQL we studied  (Figure \ref{fig:mysql-timeline}) saw the first releases to be licensed under the GPL. 
Version 3.23.31 (January, 2001) was the production release (non-beta), and the time-line view shows a flurry of topics labelled with \emph{functionality} and
\emph{maintainability}. 
After this point, this version enters the maintenance phase of its life-cycle. 
In May 2001, there is an increase in the number of topics labelled with \emph{portability}. 
This might be related to release 3.23.38, which focused on Windows compatibility. 
Similarly, in August, 2002, both \emph{functionality} and \emph{portability} are frequent, and mailing list data suggests this is related to the release
of version 3.23.52, a general bug fix with a focus on security (a component of the \emph{functionality} NFR in the ISO9126 model). 
After this point, efforts shift to the newer releases (4.0, 4.1, 5.0) and subsequently becomes more functionality oriented.

@@

In Figure \ref{fig:authorcluster} we clustered the developers from
PostgreSQL using the
Ward method of Hierarchical clustering on the 7-dimensional space.  
This will group developers according to the Euclidean distance metric.
%We chose cosine distance because it is often used in IR research to
%compare documents, in this case we are comparing the labels of topics
%associated with each author's commits.
The height in Figure \ref{fig:authorcluster} describes the euclidean
distance between elements, in terms of the path 
We clustered the authors into both $2$-clusters
and $6$-clusters. We chose $6$-clusters because the difference in distance
between entities and centroids was minimal at $6$-clusters, and the
derivative of distance was the highest in that region as well.

@@

Figure \ref{fig:authorcluster} shows that when we use $6$-clusters, 
\texttt{petere}, one of the major contributors, is in his own cluster
while two other major contributors, \texttt{tgl} and \texttt{momjian}
form their own cluster (i.e., they are distant from one another in the
7-dimensional space). 
%NEIL - not sure the relevance of this ... needs more explaining. --> (they are also the most frequent committers). 
This is interesting because it means that the important developers are
have different proportions of NFR-relevant contributions; they have different
focuses during development. In terms of the 2-clusters, we can see that \texttt{momjian}
and \texttt{tgl} are in the same cluster, but \texttt{petere} is
not. The most frequent committers do not share the same clusters, even
at the coarsest level of clustering, 2 clusters.  This implies that developers in
this sample do work on different sets of NFRs and have different
software quality focuses.

@@

Based on Figure \ref{fig:authorsim} we wanted to see which clusters
\texttt{momjian}, \texttt{tgl}, and \texttt{petere} would be in if we
evaluated the clusters proportionately, that is if we use distance
measures that normalize the data and ignore magnitude (e.g. turning an
author's NFR contributions into a histogram or unit-vector) such as
cosine distance, Euclidean distance of unit vectors, or Pearson
correlation distance ($1 - r$). Whereas, Euclidean distance uses both angle and magnitude of a
vector. For all three non-magnitude distances measures we tend to see that
the three largest contributors, \texttt{momjian}, \texttt{tgl}, and
\texttt{petere} inhabit the same clusters when we use 2-clusters or 6-clusters. This confirms the observations in
Figure \ref{fig:authorsim}, which uses Pearson correlation, that the larger contributors tend to be
proportionally similar.

One potential confound for this analysis is that we could be
describing developer style instead of developer focus. Developer style
would be a developer's likelihood to use terms found within
our dictionaries and training corpus in their commit messages. Furthermore, a top developer, in
terms of number of commits, will have more samples, and thus be more likely
take on a more general role. For example, we found in a previous
project~\cite{largechanges} that there was a good correlation between the words used to describe a commit
and the author of the commit. 
%were often correlated. Thus based
%on that observation developer style might be a confounding concern.

>>>>

% migod
In general, the fruits of mining software artifacts are often project
specific and hard to generalize.  
However, in our previous work we
investigated \emph{topic trends}---that is, topics that recur over
time---where we observed that topic trends often corresponded to
\emph{non-functional requirements} (NFRs)~\cite{Hindle09ICSM}, which is
further emphasized in this paper due to the large numbers of NFR
labelled topics.  
This is encouraging, as NFRs have the property of being cross-domain
and widely applicable. 
In this sense, they are useful abstractions for developer
conversations about different software projects.  
Furthermore, there is a series of standards on NFRs, such as ISO9126 \cite{iso9126}, that are specifically intended to apply to projects of varying
types; this suggests that our goal of trying to extract NFR-related development topics, such as those related to software quality models, holds promise.

@@

%We first introduces labelled topic extraction, and then uses that technique to analyze development activity in two large-scale examples.
We begin by discussing related work in Section~\ref{sec:related}.
Next, we describe how we generated our data (Section~\ref{sec:wordlist}). For semi-unsupervised classification (Section~\ref{sec:unsuplabelling}), we
begin by creating word-lists to signify when a topic matches an NFR label. We then apply our classifier and analyze the results. 
%and how we created the word-lists used in unsupervised classification. 
%We show that labels with their topics can be learned and used to classify other data-sets, either without training 
%Our goal is to label a given topic with either an NFR from our taxonomy,
 %Section \ref{sec:unsuplabelling} presents our technique for using word-lists to do unsupervised classification, and describes the results.
In Section~\ref{sec:suplabelling}, we manually annotate the topics, and use those annotations as training data for supervised classification.  
%We then present visualizations of named topics and their trends over time to aid communication and analysis. 
To demonstrate an application of labelled topic extraction, we use an exploratory case study of three open source database systems to show how named
topics can be compared between projects  (Section~\ref{sec:analysis}). 
The paper concludes with a discussion of limitations (Section~\ref{sec:limit}), and future work.

@@

With ConcernLines, Treude et al.~\cite{treude09cl} showed tag occurrence
using colour and intensity, and our plots that rely on color and intensity
have a similar look and feel to Treude et al.'s plots.
They mined
developer created 
change request tags from IBM Jazz
repositories and used these 
to analyze the evolution of a single product.
Change requests in Jazz allow for users to annotate each change
request and check-in with tags such as ``ui'', ``usability'' or
``milestone~3''. 
% make evolutionary analyses of a single product. 
The presence of a well-maintained set of tags is obviously essential to the success of this technique.

@@

This paper and our MSR 2011 paper~\cite{msr2011} are based off of the
work of Ernst and Mylopoulos~\cite{ernst10refsq} and the work of Hindle et
al.~\cite{Hindle09ICSM}.
In Ernst and Mylopoulos~\cite{ernst10refsq}, we describe an earlier
project that identifies changes in quality requirements
in GNOME software projects (GNOME is a Linux desktop environment). Unlike this paper, this earlier approach was more exploratory and had
less validation. In particular, it  uses different word-lists, solely uses text-matching,
and does not leverage machine learning strategies.  
Our approach in this paper, and our
MSR 2011 paper~\cite{msr2011}, extends Ernst and Mylopoulos~\cite{ernst10refsq} by using 
word-lists to label topics, which are completely absent in the earlier work.
 Hindle et
al.~\cite{Hindle09ICSM}
 propose a windowed method of topic
analysis that we extend with labelled
topics, NFRs and new visualizations. This windowed method was
to bucket documents  by time windows (such as months), that could overlap if
necessary, and then use LDA to extract documents from these
buckets. These topics were then related to topics in other adjacent
buckets, and joined if deemed sufficiently similar. This paper, and our
MSR'2011 paper~\cite{msr2011}, extend the method of Hindle et
al.~\cite{Hindle09ICSM} by labelling topics
and providing an alternative visualization, the topic time-line
described in Section \ref{sec:analysis}.

@@

For each project, we studied source control commit comments, the messages
that programmers write when they commit revisions to a source control
repository. 
Most commits we observed had comments: 90\% in MySQL 3.23,
98.5\% in PostgreSQL and 99.99\% in MaxDB 7.500.
Commit comments are often studied by researchers, as they are the most readily accessible source of project interactions, and developers are often
required to create them by the repository mechanism (e.g., CVS).  Additionally, relying only on commit comments makes our approach more generalizable,
as we do not assume the presence of other artifact corpora.
%This is the most readily accessible source of project interactions for outside researchers, since developers often or always (e.g., Git) write such
%message. Our approach is more generalizable if it does not assume the presence of other corpora. 
%bug trackers and email-list archives were not available for both projects. 
An example of a typical commit message, from MySQL, is: \textit{``history annotate diffs bug fixed (if mysql\-\_real\-\_connect() failed there were
two pointers to malloc'ed strings, with memory corruption on free(), of course)''}. 
We extracted these messages and indexed them by creation time. 
Each word in the message was
stripped of punctuation and converted to lower-case.
We summarized each message as a word distribution minus stop-words
such as ``\emph{the}'' and ``\emph{at}''. We did not stem or apply any
other transforms to the messages.
Our stop words are derived from the Natural Language Toolkit (NLTK) English stop-word
list~\footnote{NLTK: \url{http://www.nltk.org/}}.

@@

%For MySQL (Figure \ref{fig:mysql-timeline}), we similarly validated our NFR patterns with external mailing list data. 
The development period of MySQL we studied  (Figure \ref{fig:mysql-timeline}) saw the first releases to be licensed under the GPL. 
Version 3.23.31 (January, 2001) was the production release (non-beta), and the time-line view shows a flurry of topics labelled with \emph{functionality} and
\emph{maintainability}. 
After this point, this version enters the maintenance phase of its life-cycle. 
In May 2001, there is an increase in the number of topics labelled with \emph{portability}. 
This might be related to release 3.23.38, which focused on Windows compatibility. 
Similarly, in August 2002, both \emph{functionality} and \emph{portability} are frequent, and mailing list data suggests this is related to the release
of version 3.23.52, a general bug fix with a focus on security (a component of the \emph{functionality} NFR in the ISO9126 model). 
After this point, efforts shift to the newer releases (4.0, 4.1, 5.0) and subsequently becomes more functionality oriented.

@@

In Figure \ref{fig:authorcluster} we clustered the developers from
PostgreSQL using the
Ward method of Hierarchical clustering on the 7-dimensional space.  
This method groups developers according to the Euclidean distance metric.
%We chose cosine distance because it is often used in IR research to
%compare documents, in this case we are comparing the labels of topics
%associated with each author's commits.
The height in Figure \ref{fig:authorcluster} describes the euclidean
distance between elements, in terms of the path. 
We clustered the authors into both $2$-clusters
and $6$-clusters. We chose $6$-clusters because the difference in distance
between entities and centroids was minimal at $6$-clusters, and the
derivative of distance was the highest in that region as well.

@@

Figure \ref{fig:authorcluster} shows that when we use $6$-clusters, 
\texttt{petere}, one of the major contributors, is in his own cluster
while two other major contributors, \texttt{tgl} and \texttt{momjian}
form their own cluster (i.e., they are distant from one another in the
7-dimensional space). 
%NEIL - not sure the relevance of this ... needs more explaining. --> (they are also the most frequent committers). 
This is interesting because it means that the important developers
have different proportions of NFR-relevant contributions; they have different
focuses during development. In terms of the 2-clusters, we can see that \texttt{momjian}
and \texttt{tgl} are in the same cluster, but \texttt{petere} is
not. The most frequent committers do not share the same clusters, even
at the coarsest level of clustering, 2 clusters.  This implies that developers in
this sample do work on different sets of NFRs and have different
software quality focuses.

@@

Based on Figure \ref{fig:authorsim} we wanted to see which clusters
\texttt{momjian}, \texttt{tgl}, and \texttt{petere} would be in if we
evaluated the clusters proportionately, that is, if we use distance
measures that normalize the data and ignore magnitude (e.g. turning an
author's NFR contributions into a histogram or unit-vector) such as
cosine distance, Euclidean distance of unit vectors, or Pearson
correlation distance ($1 - r$). 
%Whereas Euclidean distance uses both angle and magnitude of a vector. 
For all three non-magnitude distances measures we tend to see that
the three largest contributors, \texttt{momjian}, \texttt{tgl}, and
\texttt{petere} inhabit the same clusters when we use 2-clusters or 6-clusters. This confirms the observations in
Figure \ref{fig:authorsim}, which uses Pearson correlation, that the larger contributors tend to be
proportionally similar.

One potential confound for this analysis is that we could be
describing developer style instead of developer focus. Developer style
would be a developer's likelihood to use terms found within
our dictionaries and training corpus in their commit messages. Furthermore, a top developer, in
terms of number of commits, will have more samples, and thus be more likely to 
take on a more general role. For example, we found in a previous
project~\cite{largechanges} that there was a good correlation between the words used to describe a commit
and the author of the commit. 
%were often correlated. Thus based
%on that observation developer style might be a confounding concern.
